Combination of field and experimental data together with computational models

### 2021-12-21 manuscript_1.0/manuscript_1.0.Rmd
Added a version of the manuscript in the Royal society format

---
title: "Cognitive and ecological aspects determine cleaner performance"
author:
- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  word_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: Cleaner_learning.bib
style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: Insert a nice abstract here.
editor_options:
  markdown:
    wrap: 72
---

@@

Cooperation among individuals of the same or different species is
ubiquitous in nature. Among individuals of different species it
underlays the biological mechanism for fundamental ecological
interactions such as pollination, nitrogen fixation or parasite removal
[@begon_Ecology_2006]. When partners in these mutualistic interactions
are able to choose their interaction partner, their investment and the
return they obtain can be understood as a biological market
[@hammerstein_Biological_2016; @leimar_Cooperation_2010;
@noe_Biological_1994a,@noe_Biological_1995a]. Thus, we expect supply
and demand of the good and services being exchanged to determine the
action of individuals, at least in an evolutionary time-scale. However,
supply and demand conditions of a biological market, which typically
depend on the abundance of the species involved, certainly vary in time
and space. How do partners adjust their strategy in the face of this
variation? In animals an obvious general candidate mechanism for the
strategic adjustment is the cognitive machinery. However, it is not
clear which specific cognitive mechanisms allow individual to adjust
their behaviour to the varying market conditions.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011,@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Marine_2021]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load [@grutter_Relationship_1995a]. However, 
early laboratory
experiments showed that leaving causes service priority independent of
client value [@bshary_Asymmetric_2002]. The experiments consisted of a
simultaneous two-choice task, where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014].

Recent research has refined our view on partner choice and priority of
access in the *L. dimidiatus* cleaning mutualism, and as a consequence
performance in the biological market task. First, details of the task
presentation may matter; i.e. what relevant information is encoded in
the design of the two plates (colour versus size;
@wismer_Cuebased_2019), or whether the information is in the colour of
the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
whether the food is visible or hidden/released only when the subjects
made choices on a touch screen [@pretot_Factors_2016]. Second, it
appears that cleaner and client densities in nature affect how cleaners
perform in the original market task: when densities are low, cleaners
are more likely to fail to complete the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The apparent explanation
for this failure is that visitor clients are less likely to exert their
choice options (and hence wait for inspection if made to wait) when
cleaner fish densities are low [@triki_Brain_2020]. This change in
behaviour may be due to the fact that when cleaner densities are low,
the cost of switching in terms of distance to swim and time invested
increases. If visitors are not choosy, cleaners should not give them
priority anymore. Interestingly, while cleaners from high-density areas
have larger forebrains harbouring more cells than cleaners from
low-density areas [@triki_Brain_2020], on a local scale those
individuals that perform in the market task in a way that corresponds to
the locally optimal strategy (solving/failing the task when living in
high/low density) have relatively larger forebrains [@triki_Brain_2020].
Taken together, empirical results clearly show major intra- and
inter-specific variation in performance in the market task, warranting
investigation of the underlying cognitive machinery, from perception to
decision-making [@shettleworth_Cognition_2009].

One approach to identify candidate mechanisms underlying performance in
a cognitive task is to develop proximate learning models and to evaluate
their predictions in light of the behavioural data (e.g.
@dubois_Model_2021). By varying the cognitive tool kit, one can assess
which minimal kit is necessary to solve the task at hand. This approach
has been used to analyse the market task [@prat_Role_2021;
@quinones_Reinforcement_2019]. Learning models showed that basic
reinforcement learning does not suffice to solve the task, this is
particularly so when models assume the more complex natural situation in
which cleaners not only face resident-visitor pairs but also
visitor-visitor and resident-resident pairs, as well only a resident or
only a visitor. To be able to give visitors priority over residents,
cleaners first need to be able to assess a client's value separately for
the three possible forms of encounter (alone, paired with a fish with
the same strategic option, paired with a fish with the alternative
strategic option) [@quinones_Reinforcement_2019]. In addition, cleaners
need to account for the future consequences of current decisions. In the
model, this could be achieved in two non-mutually exclusive ways
[@quinones_Reinforcement_2019]: through low temporal discounting of
future effects and/or through perceiving a visitor client leaving as
psychological punishment (i.e. as a negative reinforcer). Given that a
negative reinforcer is part of the ubiquitous learning mechanism termed
operant conditioning [@thorndike_Animal_1898; @skinner_Behavior_1938], 
this explanation can be
considered the simpler one. Furthermore, accounting for the future
effects of an action is considered an advanced cognitive ability
[@suddendorf_Evolution_2007]. Here, we fitted our model to the empirical
data on fish densities and cleaner fish performance in the market task
with the aim to determine the likely mechanism used by cleaners to
incorporate future consequences of current decisions.

@@

We used the data from @triki_Biological_2019 to fit the parameters of
the model analysed in @quinones_Reinforcement_2019. For the bayesian
fitting procedure we used Markov Chain Monte Carlo (MCMC) with a
Metropolis sampler to estimate the posterior distribution of the model
parameters. The data consist a of the number of times that a set of
cleaners chose to feed of a visitor client plate in the context of the
market task. This set of cleaners were collected in different locations
and in each one of those locations environmental variables were
measured. These environmental variables recorded from each location are
used as input for the computational model. Given these inputs, the
computational model produces a time series of choices in the market
task. In the model the preference for the two options changes as the
algorithm learns based on the reward that the two options provide. Thus,
we run the model for 10000 steps and use the last 3000 steps as the
preference developed in the model for visitors given the environmental
variables. We use this preference as the predicted probability in a
binomial distribution to calculate the likelihood that the number of
visitor choices obtained from each cleaner is generated by the model. We
sum over all the cleaners the log-likehoods and use the total for the
acceptance of new parameters in the MCMC.

@@

The model consist of a set individual-based simulations where
individuals face a series of choices between two options. Individuals in
the model represent cleaners, and the options they face simulate the
natural conditions of a cleaning market. Thus, individuals in the model
choose to prioritize different clients. These clients are classified as
resident and visitors. Residents are those clients that stay in the
cleaning queue when they are not given priority; while visitors leave
the queue (with a certain probability) after they are not given
priority. Individuals obtain a fixed reward from cleaning a client
regardless of the type. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expectation of reward and the actual
reward the individual obtains (prediction error)
[@sutton_Reinforcement_2018; \@rescola_Theory_1972]; and it is carried
in the direction that would lead to more reward being obtained, given
the new information. Further details of the model implementation can be
found in the paper [@quinones_Reinforcement_2019]. The model analysis
points to the relevance of two parameters. First, $\gamma$ measures how
much individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards in action.
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor clients as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices. The model analysis by
@quinones_Reinforcement_2019 shows that without either of them, learning
individuals cannot develop a preference for the visitor client. The
analysis contrast the effect of these two parameters against variation
in the relative abundance of the two client types. The analysis shows
that under very high client (relative) abundance the presence of future
rewards ($\gamma>0$) no longer favours the preference for the ephemeral
clients, this effect does not occur with negative reward mediated by
client behaviour ($\eta$). Thus, the analysis suggests that future
reward as an explanation for cleaner ability to prefer visitors is
supported by @triki_Biological_2019 data. However, a combination of
future and negative rewards, could still be a better explanation for the
data. In the current analysis, we will use the data from
@triki_Biological_2019 to fit some of the model parameters, particularly
relevant we will fit $\gamma$ and $\eta$ to find out whether each or a
combination of these effects is a better explanation for the pattern
seen in the data. For other parameter values used for the model
simulations and the MCMC see Table \ref{tab:param}

@@

The cognitive mechanism that our analysis highlights does not require a
direct flow of information between the cleaner and its social
environment. The adaptive responses to local conditions in cleaners is
not only evidenced by their performance in experiments, but also by
changes in their brain anatomy. Indeed, individual cleaners that
performed in the laboratory test according to what is best at the
location in which they were caught had larger forebrains harbouring more
cells than individuals whose performance mismatched locally optimal
decision-rules [@triki_Brain_2020]. This suggests that despite the
overall behavioural adjustment, evidenced here and elsewhere
[@triki_Biological_2019; @triki_Decrease_2018], there are some
cleaners better than others at adjusting their behaviour.
@triki_Brain_2020 refer to these as socially competent cleaners. Social
competence refers to the ability to optimise social behaviour depending
on the available social information [@taborsky_Social_2012a;
@bshary_Cooperation_2015]. However, we have shown here that cleaners
are able to adjust their preference according to the social environment
without using any direct information about the social environment.
Instead cleaners seem to adjust their preference using information from
the visitation patterns of their clients. These patters are of course
determined by the feedback with the cleaner's social environment
(relative abundance). Our models, however, do not include the
possibility of cleaners responding to their relative abundance directly.
The mechanism we present is a general way to learn from the environment
in both asocial and social situations. It would be interesting to
contrast this learning mechanism with one tailored to a social
situation, and see whether such mechanism fits the data better. More
generally, an important challenge ahead is to understand which cognitive
mechanisms evolved exclusively for social life.

One potential difference between cleaners that adjusted adaptively, like
the ones reported in @triki_Brain_2020, is configurational learning or
chunking. The original analysis of the learning model showed that agents
needed two mechanism to solve the market task; to account for the
future, and to assign value to the combination of clients present in the
station [@quinones_Reinforcement_2019]. The latter is referred to as
configurational learning or chunking [@sutherland_Configural_1989;
@miller_Magical_1956]. Theoretical models suggest that the lack of
configurational learning is related to a low frequency with which
stimuli are encountered together [@kolodny_Evolution_2014;
@prat_Role_2021]. Thus, cleaner relative abundance could be related to
this mechanisms as well. Given that the lack of either of these two
mechanisms yields neutral preferences [@quinones_Reinforcement_2019],
we can't, based on the current analysis, distinguish between the two
possibilities. One potential way to distinguish between the two would be
to use an experimental set up designed to test configuration learning on
cleaners with high and low performance in the market task
[@woodbury_Learning_1943; @sutherland_Configural_1989].

Combining mechanistic modelling with empirical data provides a powerful
tool to identify the cognitive processes underlying learning in animals.
The combination overcomes the limitations of the general philosophy in
animal cognition to apply the logic of Lloyd Morgen's canon (Occam's
razor). Excluding basic reinforcement learning explanations (operant
and/or classical conditioning) as a potential explanation in experiments
often warrants animals solving the task on the first possible occasion.
For example, any theory of mind task needs to be solved in the first
trial in order to exclude fast conditioning [@heyes_Theory_1998].
Similarly, subjects need to solve a social learning task on the first
trial to accept imitation as a mechanism over stimulus/local
enhancement. Such strict conditions are virtually never met. For
example, potato washing by Japanese macaques, an iconic example of
social learning, took several years to spread within the group
[@kawamura_Process_1959], meaning that any learner had been repeatedly
exposed to demonstrations before acquisition. Importantly,
@galef_Question_1992 refuted imitation as mechanism not simply because
of the repeated exposure but because a (rather qualitative) analysis of
the spread of potato washing across individuals did not follow the
prediction based on imitation learning (see also
@hirata_SweetPotato_2001). In our case, the number of trials it took
cleaners to learn the solution to the market task would never allow to
exclude an important role of negative reinforcement based on the data
alone. However, fitting model predictions to our data set revealed that
a more complex mechanism (estimation of future reward) fits the data
better.

## Cited literature

::: {#refs}
:::

@@

```{=tex}
\beginsupplement
```
```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","MCMCclean_nRew_sca_"),
                                              full.names = TRUE),value = TRUE)
knitr::include_graphics(fileNogamma)

>>>>

---
title: Combination of field and experimental data together with computational models 
  reveal cognitive mechanisms behind cleaner fish behaviour
#Cognitive and ecological aspects determine cleaner performance
author:
  - name: Andrés E. Quiñones 
    affiliation: "1"
  - name: Zegni Triki
    affiliation: "2"
  - name: Redouan Bshary
    affiliation: "1"
address:
  - code: "1"
    address: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
  - code: "2"
    address: Department of Zoology, Stockholm University, Stockholm, Sweden

corresp_author_name:  "AE. Quiñones"
corresp_author_email: "andreseqp@gmail.com"

subject:
  - "Behavioural ecology"
  - "Cognitive ecology"
  - "Animal behaviour"

keywords:
  - learning   
  - cleaners
  - behaviour
  - bayesian statisitics
  - mutualism

abstract: |
  The same behaviour can be achieved by different mechanistic underpinnings. But the environmental conditions under which the behaviour is triggered might differ between these potential mechanisms. Thus, by analyzing the varitation of expression of a certain behavior across environmental gradients, it is possible to infer which mechanism is behind the behaviour of interest. To do this it is necessary to have clearcut predictions for the different candidate mechanisms. In here, we use experimental data on the behaviour of the cleaner fish *Labroides* *dimidiatus*, as well as enviromental variables from the location of fish, together with a computational model of the learning mechanisms. Using bayesian statistics, we fit key parameters of the computational model that determine the specific learning mechanism. We find that a model where cleaners estimate the long-term consequences of their actions, and use the estimate in their desicion making, explains better under which conditions the fish develop a preference for an ephemeral food source in the context of a biological market. We argue that the combination of computational models with data is powerful tool to infer the mechanistic underpinning behind behaviours of interest.

## Remove this if not required
ethics: |
  The Animal Ethics Committee of the Queensland government (DAFF)
  approved the project (CA 2016/05/970 and CA 2017/05/1063).

data_accessibility: |
  Please provide details on the data availability.

author_contributions: |
  AQ and RB designed the study. ZT collected the data. AQ developed the model and 
  fitted the model parameters to the data. AQ, RB and ZT wrote the manuscript.

## Remove this if not required
conflict_of_interest: |
  The authors declate no conflict of interests.

## Remove this if not required
funding: |
  This work was supported by the Swiss National Science Foundation (grants
  number: 31003A_153067/1 and 310030B_173334/1 to R.B.).

## Remove this if not required
acknowledgements: |
  ZT and RB kindly thank the staff of Lizard Island Research Station, Dominique
  Roche, Sharon Wismer, Olivia Rey,Sandra Ann Binning, Elena Levorato and William McNeely for their
  field support. AQ thanks Florian Hartig for statistical advice.

bibliography: Cleanerlearning.bib

## change to true to add optional line numbering
lineno: true

header-includes: |
  ```{=latex}
  \usepackage{float}
  \usepackage{booktabs}
  \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
site: bookdown::bookdown_site
output: 
  bookdown::pdf_book:
    base_format: rticles::rsos_article
---

@@

Cooperation among individuals of the same or different species is
ubiquitous in nature. Among individuals of different species it
underlays the biological mechanism for fundamental ecological
interactions such as pollination, nitrogen fixation or parasite removal
[@begon_Ecology_2006]. When partners in these mutualistic interactions
are able to choose their interaction partner, their investment and the
return they obtain can be understood as a biological market
[@hammerstein_Biological_2016; @leimar_Cooperation_2010;
@noe_Biological_1994a;@noe_Biological_1995a]. Thus, we expect supply
and demand of the good and services being exchanged to determine the
action of individuals, at least in an evolutionary time-scale. However,
supply and demand conditions of a biological market, which typically
depend on the abundance of the species involved, certainly vary in time
and space. How do partners adjust their strategy in the face of this
variation? In animals an obvious general candidate mechanism for the
strategic adjustment is the cognitive machinery. However, it is not
clear which specific cognitive mechanisms allow individual to adjust
their behaviour to the varying market conditions.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Marine_2021]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load [@grutter_Relationship_1995a]. However, 
early laboratory
experiments showed that leaving causes service priority independent of
client value [@bshary_Asymmetric_2002]. The experiments consisted of a
simultaneous two-choice task, where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014].

Recent research has refined our view on partner choice and priority of
access in the *L. dimidiatus* cleaning mutualism, and as a consequence
performance in the biological market task. First, details of the task
presentation may matter; i.e. what relevant information is encoded in
the design of the two plates (colour versus size;
[@wismer_Cuebased_2019]), or whether the information is in the colour of
the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
whether the food is visible or hidden/released only when the subjects
made choices on a touch screen [@pretot_Factors_2016]. Second, it
appears that cleaner and client densities in nature affect how cleaners
perform in the original market task: when densities are low, cleaners
are more likely to fail to complete the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The apparent explanation
for this failure is that visitor clients are less likely to exert their
choice options (and hence wait for inspection if made to wait) when
cleaner fish densities are low [@triki_Brain_2020]. This change in
behaviour may be due to the fact that when cleaner densities are low,
the cost of switching in terms of distance to swim and time invested
increases. If visitors are not choosy, cleaners should not give them
priority anymore. Interestingly, while cleaners from high-density areas
have larger forebrains harbouring more cells than cleaners from
low-density areas [@triki_Brain_2020], on a local scale those
individuals that perform in the market task in a way that corresponds to
the locally optimal strategy (solving/failing the task when living in
high/low density) have relatively larger forebrains [@triki_Brain_2020].
Taken together, empirical results clearly show major intra- and
inter-specific variation in performance in the market task, warranting
investigation of the underlying cognitive machinery, from perception to
decision-making [@shettleworth_Cognition_2009].

One approach to identify candidate mechanisms underlying performance in
a cognitive task is to develop proximate learning models and to evaluate
their predictions in light of the behavioural data (e.g.
[@dubois_Model_2021]). By varying the cognitive tool kit, one can assess
which minimal kit is necessary to solve the task at hand. This approach
has been used to analyse the market task [@prat_Role_2021;
@quinones_Reinforcement_2019]. Learning models showed that basic
reinforcement learning does not suffice to solve the task, this is
particularly so when models assume the more complex natural situation in
which cleaners not only face resident-visitor pairs but also
visitor-visitor and resident-resident pairs, as well only a resident or
only a visitor. To be able to give visitors priority over residents,
cleaners first need to be able to assess a client's value separately for
the three possible forms of encounter (alone, paired with a fish with
the same strategic option, paired with a fish with the alternative
strategic option) [@quinones_Reinforcement_2019]. In addition, cleaners
need to account for the future consequences of current decisions. In the
model, this could be achieved in two non-mutually exclusive ways
[@quinones_Reinforcement_2019]: through low temporal discounting of
future effects and/or through perceiving a visitor client leaving as
psychological punishment (i.e. as a negative reinforcer). Given that a
negative reinforcer is part of the ubiquitous learning mechanism termed
operant conditioning [@thorndike_Animal_1898; @skinner_Behavior_1938], 
this explanation can be
considered the simpler one. Furthermore, accounting for the future
effects of an action is considered an advanced cognitive ability
[@suddendorf_Evolution_2007]. Here, we fitted our model to the empirical
data on fish densities and cleaner fish performance in the market task
with the aim to determine the likely mechanism used by cleaners to
incorporate future consequences of current decisions.

@@

We used the data from [@triki_Biological_2019] to fit the parameters of
the model analysed in [@quinones_Reinforcement_2019]. For the bayesian
fitting procedure we used Markov Chain Monte Carlo (MCMC) with a
Metropolis sampler to estimate the posterior distribution of the model
parameters. The data consist a of the number of times that a set of
cleaners chose to feed of a visitor client plate in the context of the
market task. This set of cleaners were collected in different locations
and in each one of those locations environmental variables were
measured. These environmental variables recorded from each location are
used as input for the computational model. Given these inputs, the
computational model produces a time series of choices in the market
task. In the model the preference for the two options changes as the
algorithm learns based on the reward that the two options provide. Thus,
we run the model for 10000 steps and use the last 3000 steps as the
preference developed in the model for visitors given the environmental
variables. We use this preference as the predicted probability in a
binomial distribution to calculate the likelihood that the number of
visitor choices obtained from each cleaner is generated by the model. We
sum over all the cleaners the log-likehoods and use the total for the
acceptance of new parameters in the MCMC.

@@

The model consist of a set individual-based simulations where
individuals face a series of choices between two options. Individuals in
the model represent cleaners, and the options they face simulate the
natural conditions of a cleaning market. Thus, individuals in the model
choose to prioritize different clients. These clients are classified as
resident and visitors. Residents are those clients that stay in the
cleaning queue when they are not given priority; while visitors leave
the queue (with a certain probability) after they are not given
priority. Individuals obtain a fixed reward from cleaning a client
regardless of the type. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expectation of reward and the actual
reward the individual obtains (prediction error)
[@sutton_Reinforcement_2018; @rescorla_Theory_1972]; and it is carried
in the direction that would lead to more reward being obtained, given
the new information. Further details of the model implementation can be
found in the paper [@quinones_Reinforcement_2019]. The model analysis
points to the relevance of two parameters. First, $\gamma$ measures how
much individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards in action.
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor clients as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices. The model analysis by Quiñones *et al*
[@quinones_Reinforcement_2019] shows that without either of them, learning
individuals cannot develop a preference for the visitor client. The
analysis contrast the effect of these two parameters against variation
in the relative abundance of the two client types. The analysis shows
that under very high client (relative) abundance the presence of future
rewards ($\gamma>0$) no longer favours the preference for the ephemeral
clients, this effect does not occur with negative reward mediated by
client behaviour ($\eta$). Thus, the analysis suggests that future
reward as an explanation for cleaner ability to prefer visitors is
supported by Triki *et al* [@triki_Biological_2019] data. However, a combination of
future and negative rewards, could still be a better explanation for the
data. In the current analysis, we will use the data from
[@triki_Biological_2019] to fit some of the model parameters, particularly
relevant we will fit $\gamma$ and $\eta$ to find out whether each or a
combination of these effects is a better explanation for the pattern
seen in the data. For other parameter values used for the model
simulations and the MCMC see Table \ref{tab:param}

@@

The cognitive mechanism that our analysis highlights does not require a
direct flow of information between the cleaner and its social
environment. The adaptive responses to local conditions in cleaners is
not only evidenced by their performance in experiments, but also by
changes in their brain anatomy. Indeed, individual cleaners that
performed in the laboratory test according to what is best at the
location in which they were caught had larger forebrains harbouring more
cells than individuals whose performance mismatched locally optimal
decision-rules [@triki_Brain_2020]. This suggests that despite the
overall behavioural adjustment, evidenced here and elsewhere
[@triki_Biological_2019; @triki_Decrease_2018], there are some
cleaners better than others at adjusting their behaviour. Triki *et al.*
[@triki_Brain_2020] refer to these as socially competent cleaners. Social
competence refers to the ability to optimise social behaviour depending
on the available social information [@taborsky_Social_2012a;
@bshary_Cooperation_2015]. However, we have shown here that cleaners
are able to adjust their preference according to the social environment
without using any direct information about the social environment.
Instead cleaners seem to adjust their preference using information from
the visitation patterns of their clients. These patters are of course
determined by the feedback with the cleaner's social environment
(relative abundance). Our models, however, do not include the
possibility of cleaners responding to their relative abundance directly.
The mechanism we present is a general way to learn from the environment
in both asocial and social situations. It would be interesting to
contrast this learning mechanism with one tailored to a social
situation, and see whether such mechanism fits the data better. More
generally, an important challenge ahead is to understand which cognitive
mechanisms evolved exclusively for social life.

One potential difference between cleaners that adjusted adaptively, like
the ones reported in Triki *et al.* [@triki_Brain_2020], is configurational learning or
chunking. The original analysis of the learning model showed that agents
needed two mechanism to solve the market task; to account for the
future, and to assign value to the combination of clients present in the
station [@quinones_Reinforcement_2019]. The latter is referred to as
configurational learning or chunking [@sutherland_Configural_1989;
@miller_Magical_1956]. Theoretical models suggest that the lack of
configurational learning is related to a low frequency with which
stimuli are encountered together [@kolodny_Evolution_2014;
@prat_Role_2021]. Thus, cleaner relative abundance could be related to
this mechanisms as well. Given that the lack of either of these two
mechanisms yields neutral preferences [@quinones_Reinforcement_2019],
we can't, based on the current analysis, distinguish between the two
possibilities. One potential way to distinguish between the two would be
to use an experimental set up designed to test configuration learning on
cleaners with high and low performance in the market task
[@woodbury_Learning_1943; @sutherland_Configural_1989].

Combining mechanistic modelling with empirical data provides a powerful
tool to identify the cognitive processes underlying learning in animals.
The combination overcomes the limitations of the general philosophy in
animal cognition to apply the logic of Lloyd Morgen's canon (Occam's
razor). Excluding basic reinforcement learning explanations (operant
and/or classical conditioning) as a potential explanation in experiments
often warrants animals solving the task on the first possible occasion.
For example, any theory of mind task needs to be solved in the first
trial in order to exclude fast conditioning [@heyes_Theory_1998].
Similarly, subjects need to solve a social learning task on the first
trial to accept imitation as a mechanism over stimulus/local
enhancement. Such strict conditions are virtually never met. For
example, potato washing by Japanese macaques, an iconic example of
social learning, took several years to spread within the group
[@kawamura_Process_1959], meaning that any learner had been repeatedly
exposed to demonstrations before acquisition. Importantly, Galef
[@galef_Question_1992] refuted imitation as mechanism not simply because
of the repeated exposure but because a (rather qualitative) analysis of
the spread of potato washing across individuals did not follow the
prediction based on imitation learning (see also
[@hirata_SweetPotato_2001]). In our case, the number of trials it took
cleaners to learn the solution to the market task would never allow to
exclude an important role of negative reinforcement based on the data
alone. However, fitting model predictions to our data set revealed that
a more complex mechanism (estimation of future reward) fits the data
better.

@@

```{=tex}
\beginsupplement
```
```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%',fig.pos = 'H', out.extra = ""}
fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","MCMCclean_nRew_sca_"),
                                              full.names = TRUE),value = TRUE)
knitr::include_graphics(fileNogamma)
